TacticToe: Learning the Game of Tactics

### 2018-02-07 tactictoe_jv.tex
minor

\begin{table}[ht]
\centering\ra{1.3}
\small
\begin{tabular}{llc}
\toprule
 Technique & Parameters & Solved \\
\midrule
Tactic time out & 0.02s & 154 (56.8\%)\\
                & 0.05s (default) & 156 (56.8\%)\\
                & 0.1s & 154 (56.0\%)\\
\midrule
Orthogonalization & none & 156 \\
                  & $radius = 10$ & 156 \\
                  & $radius = 20$ (default) & 156 \\ 
                  & $radius = 40$ & 156 \\

\midrule
Abstraction       & none (default)  & 156 (\%)\\ 
                  & $theorems = 8$  & 195 (\%)\\
                  & $theorems = 16$ & 199 (72.9\%)\\
                  & $theorems = 32$ & 195 (\%)\\
\midrule
MCTS Policy & $c_{policy} = 0.4$ & 149 (\%)\\
            & $c_{policy} = 0.5$ (default) & 155 (\%)\\
            & $c_{policy} = 0.6$ & 153 (\%)\\
\midrule
MCTS Evaluation & none (default) & 156 (50\%) \\
				& $c_{exploration} = 0.25$ & 153 153 (50\%)\\ 
				& $c_{exploration} = 0.5$ & 153 155 (50\%)\\ 
				& $c_{exploration} = 1$ & 154 (50\%)\\
				& $c_{exploration} =\sqrt{2}$ & 154 154 (50\%)\\ 
				& $c_{exploration} = 2$ & 152 (50\%)\\ 
\midrule
ATP integration & none (default) & 156 (50\%)\\
                & \metis 0.1s & 125 (50\%)\\
                & \metis 0.2s & 125 (50\%)\\
                & \metis 0.4s & 125 (50\%)\\ 
                & asynchronous \eprover & 125 (50\%)\\
                & \metis 0.1s + asynchronous \eprover & 190\\
\midrule
All & best parameters (without \eprover) & \\ 
\bottomrule
\end{tabular}
\end{table}

\subsection{Increasing \tactictoe's knowledge}\label{sec:perfect_exp}

\todo{decide to keep this section or not}

We investigate what are the limits of \tactictoe and what benefits additional 
knowledge could provide. To this end, we will modify the set of features to 
include features vectors that should not be known in a regular experiment. We  
include some vectors from the recorded human proof and perform the proof 
search 
after the recording instead of before in all other experiments. This 
experiment 
is not fair and should only be compared to 
experiments in the same settings.

For clarity, we will name the set of feature vectors that can be recorded from 
the current human proof $\mathbb{H}$ and the set of vectors recorded from 
previous proofs $\mathbb{P}$ (consistent with previous notation?). We note 
$\pi_1^1$ the projection on the 
first component returning the label of a feature vector. And we define the 
function $T$ that takes a tactic and return the list of its tokens.

We will progressively add bigger and bigger 
subset of $\mathbb{H}$ to $\mathbb{P}$ and observe the effect on the proof 
search success rates.

Here are the list of increasing non-trivial subsets of $\mathbb{H}$ we 
consider:
\begin{align*}
\mathbb{F}_1 &= \lbrace (t,g) \in \mathbb{H}\ |\ t \in \pi_1^1(\mathbb{P})  
   \rbrace \\
\mathbb{F}_2 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \rbrace\\
\mathbb{F}_3 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem} \rbrace\\
\mathbb{F}_4 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem or a term} 
\rbrace\\
\end{align*}

$\mathbb{F}_1$ is an approximation of what we could get if we had perfect 
prediction for those tactics.

\begin{table}[h]
\centering\ra{1.3}
\small
\begin{tabular}{lll}
\toprule
 & Additional features & Solved \\
\midrule
$\emptyset$   & none & 119 (43.6\%) \\ %NONEv2
$\mathbb{F}_1$& tactic prediction & 121 (44.3\%) \\ %AFTERTACv2
$\mathbb{F}_2$& token recombination  & 137 (50.2\%) \\ %AFTERTOKENv2
$\mathbb{F}_3$& theorem prediction  & 207 (75.8\%) \\ %AFTERTHMTHMv2
$\mathbb{F}_4$& term prediction & 223 (81.7\%) \\ %AFTERALLv2
$\mathbb{H}$  & all & 239 (87.5\%) \\ %AFTERSMALLv2
% $AFTERv2$       & $\mathfrak{H}$ & 245 (89.7\%)  \\
\bottomrule
\end{tabular}
\caption{\label{tab:featue_param} Effect of augmenting \tactictoe's knowledge 
with tactics from the proof of the tested theorem. The experiments was run with 
a time out of 5 seconds on an older version of \tactictoe, as can be seen in 
the low 
success rate of the
default strategy.} 
\end{table}

And we observe in Table~\ref{tab:featue_param} a 
slight increase compared to the default \tactictoe. We then included all 
vectors that contains tactics that can be rebuild from tokens of
tactics in our database. 
Our search algorithm proved two proofs less that if it had perfect prediction 
so it means that our prediction algorithm (policy) is doing a good  job at 
least in this early settings where proofs are easy. 
Suprisingly, being able to recombine tokens of tactics from the database is 
barely sufficient to prove more than half of the theorems in this dataset. The 
Therefore, additionally including in tactics theorems and terms that did not 
appear in previous tactics is essential. Creating from scratch other arguments 
may not help as much. This is why we restricted ourselves to predicting 
arguments of theorem list type and the term type during tactic synthesis. 
Finally our proof search algorithm sometimes fails even if it knows all the 
tactic necessary. The two main reasons are that either a crucial tactic 
exceeds its timeout or that the human proof is too deep to be rediscovered 
(more than 10 steps long).In all those experiments orthogonalization is 
performed with 
radius 20.

@@

The best \tactictoe parameters are chosen for the full-scale evaluation. Yet,
\eprover calls are not included in the best \tactictoe strategy for fairer 
comparison.

\begin{table}[h!]
\centering\ra{1.3}
\small
\begin{tabular}{llc}
\toprule
  & Parameters & Solved \\
\midrule
   \eprover   & & 2472 (34.5\%)\\ 
   \tactictoe & best parameters without \eprover & 4741 (66.2\%)\\
\midrule
   Total  &              & 4946 (69.0\%)\\
\bottomrule
\end{tabular}
\caption{Evaluation on 7164 top-level theorems of the \holfour standard library 
\label{tab:_param}}
\end{table}

@@

\paragraph{Self-learning}\todo{explain better or remove or move to future work}
All our feature vectors have been learned form human proofs. We now can now 
also add tactic-goal pairs that appears in final \tactictoe's proof. To prevent 
duplication of effort, orthogonalization of those 
tactics is essential to have a beneficial effect.
Since recording and re-proving are intertwined, the additional data is 
available for the next proof search.
The hope is that the algorithm will improve faster by learning from its own 
discovered proofs than from the human proof 
scripts~\cite{DBLP:conf/cade/Urban07}. Side experiments indicate that 
self-learning improves \tactictoe's success rate by less than one percent.

@@

\paragraph{Acknowledgments}\label{sect:acks}
This work has been supported by the
ERC Consolidator grant no.\ 649043 \textit{AI4REASON} and ERC starting
grant no.\ 714034 \textit{SMART}.

>>>>

%sh stats.sh MNM_AE2I
%273 203 3 67 0
%thibault@grid02:~/big$ sh stats.sh MNM_AE2
%273 203 3 67 0
%thibault@grid02:~/big$ sh stats.sh MNM_AE4
%273 201 3 69 0
%thibault@grid02:~/big$ sh stats.sh MNM_AE4I
%273 200 3 70 0
%thibault@grid02:~/big$ sh stats.sh MNM_AE1
%273 198 3 72 0
%thibault@grid02:~/big$ sh stats.sh MNM_AE1I
%273 200 3 70 0

%thibault@grid02:~/big$ sh stats.sh MNM_ATPE
%273 213 3 57 0
%thibault@grid02:~/big$ sh stats.sh MNM_ATPM01
%273 216 3 54 0
%thibault@grid02:~/big$ sh stats.sh MNM_ATPM01E
%273 218 3 52 0
%thibault@grid02:~/big$ sh stats.sh MNM_ATPM02
%273 212 3 58 0
%thibault@grid02:~/big$ sh stats.sh MNM_ATPM04
%273 212 3 58 0

%thibault@grid02:~/big$ sh ftim.sh MNM_O40
%127
%9
%8
%2
%2
%148
%thibault@grid02:~/big$ sh ftim.sh MNM_O10
%124
%15
%5
%3
%1
%148
%thibault@grid02:~/big$ sh ftim.sh MNM_ON
%108
%16
%6
%3
%6
%thibault@grid02:~/big$ sh ftim.sh MNM_T005
%124
%13
%5
%2
%3

\begin{table}[ht]
\centering\ra{1.3}
\small
\begin{tabular}{llcc}
\toprule
 Technique & Parameters & Solved (1s) & Solved (60s)\\
\midrule
Tactic time out & 0.02s && 154\\
                & 0.05s (default) && 156\\
                & 0.1s && 154\\
\midrule
Orthogonalization & none & 105 & 156 \\
                  & $radius = 10$ & 121 & 156 \\
                  & $radius = 20$ (default) & 121 & 156 \\ 
                  & $radius = 40$ & 124 & 156 \\
\midrule
Abstraction       & none (default)  && 156\\ 
                  & $theorems = 8$  && 195\\
                  & $theorems = 16$ && 199\\
                  & $theorems = 32$ && 195\\
\midrule
MCTS policy & $c_{policy} = 0.4$ && 149\\
            & $c_{policy} = 0.5$ (default) && 156\\
            & $c_{policy} = 0.6$ && 153\\

\midrule
MCTS evaluation & none (best abstraction) && 199\\
				& $c_{exploration} = 1$ && 201\\ 
				& $c_{exploration} = 2$ && 203\\ 
				& $c_{exploration} = 4$ && 198\\
\midrule
ATP integration & none (best MCTS evaluation) && 203\\
                & \metis 0.1s && 216\\
                & \metis 0.2s && 212\\
                & \metis 0.4s && 212\\ 
                & \eprover && 213\\
                & \metis 0.1s + \eprover && 218\\
\bottomrule
\end{tabular}
\end{table}

Say the respective percentage after 5seconds for orthogonalization.

%\subsection{Increasing \tactictoe's knowledge}\label{sec:perfect_exp}
%
%\todo{decide to keep this section or not}
%
%We investigate what are the limits of \tactictoe and what benefits additional 
%knowledge could provide. To this end, we will modify the set of features to 
%include features vectors that should not be known in a regular experiment. We  
%include some vectors from the recorded human proof and perform the proof 
%search 
%after the recording instead of before in all other experiments. This 
%experiment 
%is not fair and should only be compared to 
%experiments in the same settings.
%
%For clarity, we will name the set of feature vectors that can be recorded from 
%the current human proof $\mathbb{H}$ and the set of vectors recorded from 
%previous proofs $\mathbb{P}$ (consistent with previous notation?). We note 
%$\pi_1^1$ the projection on the 
%first component returning the label of a feature vector. And we define the 
%function $T$ that takes a tactic and return the list of its tokens.
%
%We will progressively add bigger and bigger 
%subset of $\mathbb{H}$ to $\mathbb{P}$ and observe the effect on the proof 
%search success rates. 
%
%Here are the list of increasing non-trivial subsets of $\mathbb{H}$ we 
%consider:
%\begin{align*}
%\mathbb{F}_1 &= \lbrace (t,g) \in \mathbb{H}\ |\ t \in \pi_1^1(\mathbb{P})  
%   \rbrace \\
%\mathbb{F}_2 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \rbrace\\
%\mathbb{F}_3 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem} \rbrace\\
%\mathbb{F}_4 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem or a term} 
%\rbrace\\
%\end{align*}
%
%
%$\mathbb{F}_1$ is an approximation of what we could get if we had perfect 
%prediction for those tactics.
%
%\begin{table}[h]
%\centering\ra{1.3}
%\small
%\begin{tabular}{lll}
%\toprule
% & Additional features & Solved \\
%\midrule
%$\emptyset$   & none & 119 (43.6\%) \\ %NONEv2
%$\mathbb{F}_1$& tactic prediction & 121 (44.3\%) \\ %AFTERTACv2
%$\mathbb{F}_2$& token recombination  & 137 (50.2\%) \\ %AFTERTOKENv2
%$\mathbb{F}_3$& theorem prediction  & 207 (75.8\%) \\ %AFTERTHMTHMv2
%$\mathbb{F}_4$& term prediction & 223 (81.7\%) \\ %AFTERALLv2
%$\mathbb{H}$  & all & 239 (87.5\%) \\ %AFTERSMALLv2
%% $AFTERv2$       & $\mathfrak{H}$ & 245 (89.7\%)  \\
%\bottomrule
%\end{tabular}
%\caption{\label{tab:featue_param} Effect of augmenting \tactictoe's knowledge 
%with tactics from the proof of the tested theorem. The experiments was run 
%with 
%a time out of 5 seconds on an older version of \tactictoe, as can be seen in 
%the low 
%success rate of the
%default strategy.} 
%\end{table}
%
%
%And we observe in Table~\ref{tab:featue_param} a 
%slight increase compared to the default \tactictoe. We then included all 
%vectors that contains tactics that can be rebuild from tokens of
%tactics in our database. 
%Our search algorithm proved two proofs less that if it had perfect prediction 
%so it means that our prediction algorithm (policy) is doing a good  job at 
%least in this early settings where proofs are easy. 
%Suprisingly, being able to recombine tokens of tactics from the database is 
%barely sufficient to prove more than half of the theorems in this dataset. The 
%Therefore, additionally including in tactics theorems and terms that did not 
%appear in previous tactics is essential. Creating from scratch other arguments 
%may not help as much. This is why we restricted ourselves to predicting 
%arguments of theorem list type and the term type during tactic synthesis. 
%Finally our proof search algorithm sometimes fails even if it knows all the 
%tactic necessary. The two main reasons are that either a crucial tactic 
%exceeds its timeout or that the human proof is too deep to be rediscovered 
%(more than 10 steps long).In all those experiments orthogonalization is 
%performed with 
%radius 20.

@@

The best \tactictoe parameters are chosen for the full-scale evaluation. Yet,
\eprover asynchronous calls are not included in the best \tactictoe strategy 
for fairer comparison.

\begin{table}[h!]
\centering\ra{1.3}
\small
\begin{tabular}{lc}
\toprule
  & Solved (60s) \\
\midrule
   \eprover   & 2472 (34.5\%)\\ 
   \tactictoe & 4741 (66.2\%)\\
\midrule
   Total  & 4946 (69.0\%)\\
\bottomrule
\end{tabular}
\caption{Evaluation on 7164 top-level theorems of the \holfour standard library 
\label{tab:_param}}
\end{table}

@@

%\paragraph{Self-learning}\todo{explain better or remove or move to future work}
%All our feature vectors have been learned form human proofs. We now can now 
%also add tactic-goal pairs that appears in final \tactictoe's proof. To 
%prevent 
%duplication of effort, orthogonalization of those 
%tactics is essential to have a beneficial effect.
%Since recording and re-proving are intertwined, the additional data is 
%available for the next proof search.
%The hope is that the algorithm will improve faster by learning from its own 
%discovered proofs than from the human proof 
%scripts~\cite{DBLP:conf/cade/Urban07}. Side experiments indicate that 
%self-learning improves \tactictoe's success rate by less than one percent.

@@

\paragraph{Acknowledgments}\label{sect:acks}
We would like to thank Lasse Blaauwbroek for his insightful comments which
contributed to improve the quality of this paper. This work has been supported 
by the ERC Consolidator grant no.\ 649043 \textit{AI4REASON} and ERC starting
grant no.\ 714034 \textit{SMART}.
